---
title: "GSS challange More Data"
author: "Genesis Alfaro"
date: "6-17-24"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
mydata <- select(GSS,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

  This is a type of categorical problem, with at least one problem with multiple levels. To account for this we will use Fisher test.
  
  
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

 In this graph, the first slide displays the number of individual categorized by race, white, black and other. The white race had the with over 1500 reports, and the black and other race had less than 250 reports. In the last display the chart shows the percentage of the races in favor of capital punishment. Each category was compared to each other, the white race was compared to support versus the opposition of capital punishment and was in favor of capital punishment by approximately 64% The black race was in opposition, with a much lower percentage of around 57%. The other race was in favor of capital punishment with a near report for its opposition, the support percentage was around 55%.
 
 
## Now with The Most data (fewer variables gives more data points)

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
The summary displays that 65% of the white race was reported, 42% of the black race, and 55% of the other race were all in favor f capital punishment. This is compared to 34% of the white race in opposition of capital punishment, 57% of the black race, ad 44% of the other race.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with More data
```{r}
fisherXY(x=mydata$race, y=mydata$cappun, printAll = "TRUE")
```



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`
```{r}
chisq.test(table2)
chisqtestGC(table2)
```


## More data version
```{r}
fisherXY(mydata$race,mydata$cappun, printALL= "TRUE")
```
In the inferential data we ran a CHI Square and fisher test to get results of each comparison. The first test, the CHI Square shared that the probability it pulled was equal to less than .000000000000000022. This number is too small.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The Fisher test confirms the comparisons of each race. The white race when compared to the black race was significantly different. The resulting p value was .00000174. The white and other race were compared and the p value that resulted was 0.2565. In the last section of comparison, the black and other races were compared and the resulting p value was 0.06299. This observation of the Fisher test is a replication of the process used in the master branch.
The p value of both CHI squared test and the Fisher test are not in favor are not in favor of the results and we now reject the null hypothesis.
