---
title: "GSS challange Less Data"
author: "Genesis Alfaro"
date: "Summer 2024"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

In this branch we are duplicating the process done in the master branch, with more data than previous by comparing similar tests and their results. This will sort through more data by using correlating variables to data points.


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
The graphs are in the same format as in the master branch. The information is targeted better with a smaller scope of observations. In the first slide there is a smaller number of white race around 400 reports in favor of capitol punishment. This number is still almost three times larger than the other two races. Both the black and other races have less than 100 reports in favor. The opposing reports are drastically different in  the white race column, with just under half of all the favorable reports.

## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.

The graphs are in the same format as in the master branch. The information s targeted better with a smaller scope of  observations. In the first slide there is a smaller number of the white race at around 400 reports in favor of capital punishment. This number is still at least three times larger than the other two races. Both the black and other races have less than 100 reports in favor. The opposing reports are different in te white race column, with just under half of all the favorable reports.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

In the numerical results we see similar numbers to the master branches results. The white race with 67% in favor of capital punishment. The black race contains 47% in favor, and the other race with 61%. This mimics the numbers of the chart above, and can be seen as well in the master branch.



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfysthe ssumptos and is appropriate to ach answer the research question!-->

```{r}
fisherXY(x=mydata$race, y=mydata$cappun, confL = .01, printAll = "TRUE") 
```

Using the CHI test and  fisher test the numerical results gives us insight into the differences between each comparison using th new data. The CHI test information provides a probability values of less than .000000000000000022%. This number is too small.
The Fisher test confirms the comparisons of each race. The white race when compared to the black race was significantly different. The resulting p value was .00000174. The white and other race were compared and the p value that resulted was 0.2565. In the last section of comparison, the black and other races were compared and the resulting p value was 0.06299. This observation of the Fisher test is a replication of the process used in the master branch.
The p value of both CHI squared test and the Fisher test are not in favor are not in favor of the results and we now reject the null hypothesis.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
